Guided Proofreading of Automatic Segmentations for Connectomics

### 2016-11-15 paper/method.tex
Section 3 review.

We build a split error classifier with output $p$ using a CNN to check whether an edge within an existing automatic segmentation is valid ($p=0$) or not ($p=1$). Rather than analyzing every input pixel, the classifier operates only on segment boundaries, which requires less pixel context and is faster. In contrast to Bogovic \etal~\cite{BogovicHJ13}, we work with 2D slices rather than 3D volumes. This enables proofreading prior or in parallel to a computationally expensive alignment of individual EM images.
\\~\\
\textbf{Convolutional Neural Network Architecture.} Split error detection of a given boundary is a binary classification task since the boundary is either correct or erroneous. However, in reality the score $p$ is between 0 and 1. The classification complexity arises from hundreds of different cell types in connectomics data rather than from the classification decision itself. Intuitively, this yields a wider architecture with more filters rather than a deeper architecture with more layers. We explored different architectural configurations--including residual networks~\cite{resnet}--by performing a brute force parameter search and comparing precision and recall (see supplementary materials). Our final CNN configuration for split error detection is composed of four convolutional layers, each followed by max pooling as well as dropout regularization to prevent overfitting due to limited training data. Fig.~\ref{fig:architecture} shows the CNN architecture for split error detection.
\\~\\
\textbf{Classifier Inputs.} To train the CNN for split error detection, we take boundary context information into consideration for the decision making process. For this, we use a $75\times75$ pixel patch at the center of an existing boundary. This covers approximately $80\%$ of all boundaries in real-world connectomics data with nanometer resolution. If the boundary length is not fully covered, we sample up to 10 non-overlapping patches along the boundary and combine the resulting score by weighted averaging based on boundary length coverage per patch.
%
Similar to Bogovich~\etal~\cite{BogovicHJ13}, we use grayscale image data,
corresponding boundary probabilities, and a single binary mask combining the two
neighboring labels as features for our CNN. However, we observed that the
boundary probability information generated from EM images is often misleading
due to noise or artifacts in the data. This can result in merge errors within
the automatic segmentation. To better direct our classifier to train on the true
boundary, we extract the border between two segments. We then dilate this border
by 5 pixels to consider slight edge ambiguities and use this binary mask as an
additional feature to create a stacked 4-channel input patch.
Fig.~\ref{fig:cnn_inputs} shows examples of correct and erroneous feature
patches and their corresponding automatic segmentation and ground truth.

\begin{figure}[h]
\begin{center}
  \includegraphics[width=\linewidth]{gfx/cnn_inputs.pdf}
\end{center}
 % \vspace{-4mm}
   \caption{Example inputs for learning correct splits and split errors as reflected in the segmentation relative to the ground truth. Image, membrane probabilities, merged binary labels, and a dilated border mask are combined to 4-channel input patches.}
\label{fig:cnn_inputs}
\end{figure}

@@

Identification and correction of merge errors is more challenging than finding and fixing split errors, because we must look inside segmentation regions for missing or incomplete boundaries and then propose the correct boundary. However, we can reuse the same trained CNN for this task. Similar to guided volume editing by Karimov~\etal~\cite{karimov_guided_volume_editing} we generate potential borders within a segment. For each segmentation label, we dilate the label by 20 pixel and generate 50 potential boundaries through the region by randomly placing watershed seed points at opposite sides of the label boundary. For watershed, we use the inverted grayscale EM image as features. This yields 50 candidate splits.

Dilation of the segment prior to watershed is motivated by our observation that the generated split tends to attach to the real membrane boundary. These boundaries are then individually rated using our split error classifier. For this, we invert the probability score such that a correct split (previously encoded as $p=0$) is most likely a candidate for a merge error (now encoded as $p=1$). In other words, if a generated boundary is ranked as correct, it probably should be in the segmentation. Fig. \ref{fig:merge_error} illustrates this procedure.

\begin{figure}[h]
\begin{center}
  \includegraphics[width=\linewidth]{gfx/merge_error.pdf}
\end{center}
%  \vspace{-4mm}
   \caption{Merge errors are identified by generating randomly seeded watershed borders within a dilated label segment. These borders then are individually rated using the split error CNN by inverting the probability score. This way, a confident rating for a correct split most likely indicates the missing border of the merge error and can be used for correcting the labeling.}
\label{fig:merge_error}
\end{figure}

@@

We use the proposed classifiers in combination to perform corrections of split and merge errors in automatic segmentations. For this, we first perform merge error detection for all existing segments in a dataset and store the inverted rankings $1-p$ as well as potential corrections. After that, we perform split error detection and store the ranking $p$ for all neighboring segments in the segmentation. We then sort the merge and split error rankings individually from highest to lowest.
For error correction, we first loop through the potential merge error regions and then through the potential split error regions. During this process, each error is now subject to a yes/no decision which can be provided in different ways:
\\~\\
\textbf{Selection oracle.} If ground truth data is available, the selection oracle \textit{knows} whether a possible correction improves an automatic segmentation. This is realized by simply comparing the outcome of a correction using a defined measure. The oracle only accepts corrections which improve the automatic segmentation---others get
discarded. This equals perfect proofreading.
\\~\\
\textbf{Automatic selection with threshold.} The decision whether to accept or reject a potential correction is done by comparing rankings to a threshold $p_t$. If the inverted score $1-p$ of a merge error is higher than a threshold $1-p_t$, the correction is accepted. Similarly, a correction is accepted for a split error if the ranking $p$ is higher than $p_t$. Our experiments have shown that the threshold $p_t$ is the same for merge and split errors for a balanced classifier that has been trained on equal numbers of correct and error patches.
\\~\\
\textbf{Forced choice setting.} A user is presented with the choices of either accepting or rejecting a correction. This way, all potential split errors are seen. Inspecting all merge errors is not possible for users due to the sheer amount of generated borders. We therefore only present merge errors that satisfy $1-p_t$.
\\~\\
In all cases, a decision has to be made to advance to the next possible erroneous region. If a merge error correction was accepted, the newly found boundary is added to the segmentation data. This partially updates the merge error and split error ranking with respect to the new segment. If a split error correction was accepted, two segments are merged in the segmentation data and the disappearing segment is removed from all error rankings. We then perform merge error detection on the now larger segment and update the ranking. We also update the split error rankings to include all new neighbors and re-sort. The error with the next highest ranking is now subject to the yes/no decision.

@@

Our guided proofreading is integrated into an existing workflow for large connectomics data. The system is web-based and is designed with a user-friendly user interface (Fig.\ref{fig:ui}) showing three components. We show the outline of the current labeling of a cell boundary and its proposed correction overlaying the EM image data. For the user, it is not possible to distinguish the current labeling and the proposed correction to avoid selection bias. We also show a solid overlay of the current and the proposed labeling. In addition, we show the image without overlays to provide an unoccluded view. User interaction is simple and involves one mouse click on either the current labeling or the correction. After interaction, the next potential error is shown.

\begin{figure}[h]
\begin{center}
  \includegraphics[width=\linewidth]{gfx/user_interface_split.pdf}
\end{center}
  \vspace{-4mm}
   \caption{Our guided proofreading user interface for Segmentation correction. An image without overlays is shown on the left. A split error (right) and its correction (center) are the possible choices for the user. Hovering highlights the current selection with a yellow border, and a mouse click confirms it by advancing to the next potential error.}
\label{fig:ui}
\end{figure}

>>>>

We build a split error classifier with output $p$ using a CNN to check whether an edge within an existing automatic segmentation is valid ($p=0$) or not ($p=1$). Rather than analyzing every input pixel, the classifier operates only on segment boundaries, which requires less pixel context and is faster. In contrast to Bogovic \etal~\cite{BogovicHJ13}, we work with 2D slices rather than 3D volumes. This enables proofreading prior or in parallel to a computationally expensive stitching and 3D alignment of individual EM images.
\\~\\
\textbf{Convolutional Neural Network Architecture.} Split error detection of a given boundary is a binary classification task since the boundary is either correct or erroneous. However, in reality, the score $p$ is between 0 and 1. The classification complexity arises from hundreds of different cell types in connectomics data rather than from the classification decision itself. Intuitively, this yields a wider architecture with more filters rather than a deeper architecture with more layers. We explored different architectural configurations---including residual networks~\cite{resnet}---by performing a brute force parameter search and comparing precision and recall (see supplementary materials). Our final CNN configuration for split error detection is composed of four convolutional layers, each followed by max pooling as well as dropout regularization to prevent overfitting due to limited training data (Fig.~\ref{fig:architecture}).
\\~\\
\textbf{Classifier Inputs.} To train the CNN, we consider boundary context in the decision making process by using a $75\times75$ pixel patch at the center of an existing boundary. This size covers approximately $80\%$ of all boundaries in the 6~nm Mouse S1 AC3 Open Connectome Project dataset. If the boundary length is not fully covered, we sample up to 10 non-overlapping patches along the boundary, and combine the resulting score by averaging weighted by the boundary length coverage per patch.

Similar to Bogovich~\etal~\cite{BogovicHJ13}, we use grayscale image data,
corresponding boundary probabilities, and a single binary mask combining the two
neighboring labels as inputs to our CNN. However, we observed that the
boundary probability information generated from EM images is often misleading
due to noise or artifacts in the data. This can result in merge errors within
the automatic segmentation. To better direct our classifier to train on the true
boundary, we extract the border between two segments. Then, we dilate this border
by 5 pixels to consider slight edge ambiguities and use this binary mask as an
additional input. This creates a stacked 4-channel input patch.
Fig.~\ref{fig:cnn_inputs} shows examples of correct and erroneous input
patches and their corresponding automatic segmentation and ground truth.

\begin{figure}[t]
\begin{center}
  \includegraphics[width=\linewidth]{gfx/cnn_inputs.pdf}
\end{center}
 % \vspace{-4mm}
   \caption{Example inputs for learning correct splits and split errors, as per a candidate segmentation versus the ground truth. Image, membrane probabilities, merged binary labels, and a dilated border mask provide 4-channel input patches.}
\label{fig:cnn_inputs}
\end{figure}

@@

Identification and correction of merge errors is more challenging than finding and fixing split errors, because we must look inside segmentation regions for missing or incomplete boundaries and then propose the correct boundary. However, we can reuse the same trained CNN for this task. Similar to guided volume editing by Karimov~\etal~\cite{karimov_guided_volume_editing}, we generate potential borders within a segment. For each segmentation label, we dilate the label by 20 pixel and generate 50 potential boundaries through the region by randomly placing watershed seed points at opposite sides of the label boundary. We perform watershed on the inverted grayscale EM image. This yields 50 candidate splits.

Dilation of the segment prior to watershed is motivated by our observation that the generated splits tend to attach to real membrane boundaries. These boundaries are then individually rated using our split error classifier. For this, we invert the probability score such that a correct split (previously encoded as $p=0$) is most likely a candidate for a merge error (now encoded as $p=1$). In other words, if a generated boundary is ranked as correct, it probably should be in the segmentation. Fig. \ref{fig:merge_error} illustrates this procedure.

\begin{figure}[t]
\begin{center}
  \includegraphics[width=\linewidth]{gfx/merge_error.pdf}
\end{center}
%  \vspace{-4mm}
   \caption{Merge errors are identified by generating randomly-seeded watershed borders within a dilated label segment. Then, each border is individually rated using the split error CNN by inverting the probability score. A confident rating for a correct split most likely indicates the missing border of the merge error, and can be used to correct the labeling.\JT{Candidate to improve.}}
\label{fig:merge_error}
\end{figure}

@@

We use the proposed classifiers in combination to perform corrections of split and merge errors in automatic segmentations. For this, we first perform merge error detection for all existing segments in a dataset and store the inverted rankings $1-p$ as well as potential corrections. After that, we perform split error detection and store the ranking $p$ for all neighboring segments in the segmentation. Then, we sort the merge and split error rankings separately from highest to lowest. For error correction, first we loop through the potential merge error regions and then through the potential split error regions. During this process, each error is now subject to a yes/no decision which can be provided in different ways:

\paragraph{Selection oracle.} If ground truth data is available, the selection oracle \textit{knows} whether a possible correction improves an automatic segmentation. This is realized by simply comparing the outcome of a correction using a defined measure. The oracle only accepts corrections which improve the automatic segmentation---others get
discarded. This is guided proofreading with a perfect user, and allows us to assess the upper limit of performance.

\paragraph{Automatic selection with threshold.} The decision whether to accept or reject a potential correction is taken by comparing rankings to a threshold $p_t$. If the inverted score $1-p$ of a merge error is higher than a threshold $1-p_t$, the correction is accepted. Similarly, a correction is accepted for a split error if the ranking $p$ is higher than $p_t$. Our experiments have shown that the threshold $p_t$ is the same for merge and split errors for a balanced classifier that has been trained on equal numbers of correct and error patches.

\paragraph{Forced choice setting.} We present a user with the choice to accept or reject a correction. All potential split errors are seen. Inspecting all merge errors is not possible for users due to the sheer amount of generated borders. Therefore, we only present merge errors that satisfy $1-p_t$.

\noindent \newline In all cases, a decision has to be made to advance to the next possible erroneous region. If a merge error correction was accepted, the newly found boundary is added to the segmentation data. This partially updates the merge error and split error ranking with respect to the new segment. If a split error correction was accepted, two segments are merged in the segmentation data and the disappearing segment is removed from all error rankings. Then, we perform merge error detection on the now larger segment and update the ranking. We also update the split error rankings to include all new neighbors, and re-sort. The error with the next highest ranking then forces a choice.

@@

Our guided proofreading is integrated into an existing workflow for large connectomics data. The system is web-based and is designed with a novice-friendly user interface (Fig.~\ref{fig:ui}). We show the outline of the current labeling of a cell boundary and its proposed correction overlaying the EM image data. For the user, it is not possible to distinguish the current labeling and the proposed correction to avoid selection bias. We also show a solid overlay of the current and the proposed labeling. In addition, we show the image without overlays to provide an unoccluded view. User interaction is simple and involves one mouse click on either the current labeling or the correction. After interaction, the next potential error is shown.

\begin{figure}[t]
\includegraphics[width=\linewidth]{gfx/user_interface_split.pdf}
\caption{Our guided proofreading user interface. A candidate error region is shown on the left. The user must choose between the region being a split error which needs correcting (center) or not (right). Hovering highlights the current selection with a yellow border, and a mouse click confirms the choice and advances to the next potential error.\JT{Needs labels on the images; also, removal of black background.}}
\label{fig:ui}
\end{figure}

### 2016-11-15 paper/related_work.tex
Section 3 review.

\begin{figure*}[t]
\begin{center}
\includegraphics[width=\linewidth]{gfx/architecture.pdf}
\end{center}
  \vspace{-4mm}
   \caption{We build the guided proofreading classifiers using a traditional CNN architecture. The network is based on four convolutional layers, each followed by max pooling as well as dropout regularization. The 4-channel input patches are rated as either correct splits or as split errors.}
\label{fig:architecture}
\end{figure*}

>>>>

\begin{figure*}[t]
\begin{center}
\includegraphics[width=\linewidth]{gfx/architecture.pdf}
\end{center}
  \vspace{-4mm}
   \caption{Our guided proofreading classifiers use a traditional CNN architecture of four convolutional layers, each followed by max pooling and dropout regularization. The 4-channel input patches are labelled either as correct splits or as split errors.}
\label{fig:architecture}
\end{figure*}
